\documentclass{book}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\begin{document}
\pagenumbering{none}
\begin{flushright}
\noindent Igor Yur'evich Popov,\\
Professor, Head of Mathematics Department of \\
Saint-Petersburg State University of \\
Information Technology, Mechanics and Optics.\\

Email: popov@mail.ifmo.ru
\end{flushright}
\vspace{20pt}
\section*{\protect \centering Letter of Recommendation}
\vspace{10pt}
To Whom it May Concern,

\vspace{10pt}
\noindent This letter of recommendation is provided to Andrey Breslav, a second-year PhD student at Mathematics Department of Saint-Petersburg State University of Information Technology, Mechanics and Optics, who asked me to serve as a reference on his behalf. Andrey is doing his research under my supervision for last two years. During this period he showed an ability to perform successful research and actively participated in teaching activities. 

Andrey works on problems connected to software engineering and currently his studies are concentrated around technology for creating textual syntax for programming languages and making language development artifacts modular, reusable and maintainable. He proposed usage of aspect-oriented techniques in grammar development and developed a system that supports these techniques. He has published a paper on this topic and given two talks at Russian conferences. Andrey's research is supported by a grant from Saint-Petersburg government.

During our joint work Andrey showed a strong theoretical basis and a will to it's continuous growth. He is capable of self-education and regularly attends international and local schools to advance his knowledge and skills. He also has industrial experience in software development.

Andrey worked as a lecturer at our department for three years. He gave the following courses for fist- and second- year students:
\begin{list}{$\bullet$}{\topsep 0pt \itemsep -3pt}
	\item Introduction to Programming 
	\item Algorithms and Data Structures
	\item Object-Oriented Programming
	\item Java Programming Language
\end{list}
He has supported his courses with original lecture notes and problem sets.
For now there are three students from different institutions working on their bachelor theses under his supervision.

As a conclusion I wold like to state my strong recommendation for Andrey Breslav.

\vspace{45pt}
\begin{tabular}{rlr}%
Igor Yur'evich Popov, & \hspace{50pt} & \today \\
Professor, Head of Mathematics Department of &&\\
Saint-Petersburg State University of &&\\
Information Technology, Mechanics and Optics &&\\
\end{tabular}%
\end{document}
